---
title: "Untitled"
author: "Jessica Hunter"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

three key functions of a function
You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.
You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).
1.You place the code you have developed in body of the function, a { block that immediately follows function(...).

2. You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.
3. You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).
4. You place the code you have developed in body of the function, a { block that immediately follows function(...).

return tells us what values we want returned when you run the function
**never create from scratch** start like regular function and then work backwards

### steps

step1. name the function




```{r}
library(tidyverse)

df <- tibble::tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)


```



```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)))

```



```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))
```


##creating my  for F to C degrees
start with function
```{r}
# temp_C <- (temp_F-32) * 5 / 9
```
name function
```{r}
# F_to_C <- function(){}
```

put in equation  
put in argumant  
what value i want returned back? temp c  
```{r}
F_to_C <- function(temp_F){
  temp_C <- (temp_F-32) * 5 / 9
  return(temp_C)
  }
```

test it
```{r}
F_to_C(32)
```

create function c to F
ˆ
```{r}
C_to_K <- function(temp_C){
  temp_K <- temp_C + 273.15
  return(temp_K)
}
```

```{r}
C_to_K(0)
```



