---
title: "Week 8 Homework"
author: "Jessica Hunter"
date: "3/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction
The purpose of this script is to create functions to view the locations of tree species in San Francisco and also see the 





## Libraries

```{r}
library(tidyverse)
library(here)
library(maps)
library(mapproj)
library(mapdata)
library(ggplot2)
library(ggmap)
```

## Load Data
Data from Tidy Tuesday January 28, 2022
```{r}
sf_trees <- read_csv(here("week_8","data8", "sf_trees.csv"))

# sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

## Functions
This section contains the functions used to find the names and plot the locations of plant species.

#### Filter data for San Francisco map

```{r}
# to make a map of san francisco
sanfran <- map_data("county") %>% 
 filter(subregion == "san francisco")

```

### Data cleaning of names of common and scientific species

```{r}
# separates the columns into the scientific or common name and cleans the data
sf_trees_clean <-sf_trees %>% 
  select(species, latitude, longitude) %>% 
  separate(col = species,
           into = c("scientific","common"),
           sep = " :: ") %>% 
  filter(common != "") %>% 
  separate(col = scientific,
           into = c("scientific", "trash"),
           sep = " '") %>% 
  separate(col = common,
           into = c("common", "othercommon"),
           sep = " '") %>% 
  separate(col = othercommon,
           into = c("othercommon"),
           sep = "'") %>% 
  separate(col = common,
           into = c("common", "othercommon2"),
           sep = ": ") %>%  
  unite(col = "othercommon",
       c(othercommon,othercommon2),
       na.rm = TRUE) %>% 
  select(scientific, common, othercommon, latitude, longitude) %>% 
  filter(longitude> -122.4812) %>% 
  drop_na(latitude)
```



### data with tree names

```{r}

left_sf_tree_names<-sf_trees_clean %>% 
  select(common, scientific)

left<-unique.data.frame(left_sf_tree_names)

right_sf_tree_names<-sf_trees_clean %>% 
  select(common=othercommon, scientific)
right<-unique.data.frame(right_sf_tree_names)

treenames<-full_join(left,right) %>% 
  filter(common != "")


```



### prints tree scientific names from common names

```{r}
common2scientific<- function(samplename, data=treenames){
  if (samplename %in% treenames$common){
    abcd<-filter(treenames, common=={{samplename}}) %>% 
    print()
  } else{
    print("make sure to use proper capitals and quotes around names")
  }
}

```



```{r}
scientific2common<- function(samplename, data=treenames){
  if (samplename %in% treenames$scientific){
    abcd<-filter(treenames, scientific=={{samplename}}) %>% 
    print()
  } else{
    print("no")
  }
}
```



### Function where you enter scientific tree name and plots on maps
```{r}
sftreeplot<- function(samplename,data=sf_trees_clean){
  if (samplename %in% sf_trees_clean$scientific){
    abcd<-filter(sf_trees_clean, scientific=={{samplename}}) %>% 
      ggplot()+
  geom_point(aes(x=longitude,y=latitude,
                 color = scientific))+
    geom_polygon(data = sanfran,
                aes(y = lat,
                    x = long,
                    group = group,
                    alpha = 9))+
  coord_map(projection = "mercator",
             ylim = c(37.69,37.825),
             xlim = c(-122.5, -122.3))+
  theme(legend.position = "none")+
      guides(fill=FALSE)
  abcd
  } else{
    print("error:Did you place in quotes?")
  }
}


```


## Analysis Examples

### List the common names for the scientific name "Ginkgo biloba"

```{r}
scientific2common("Ginkgo biloba")
```

### List the scientific name for the common name "Autumn Gold"
```{r}
common2scientific("Autumn Gold")
```

### Plot the locations of tree species "Ginkgo biloba" in San Francisco County

```{r}
sftreeplot("Ginkgo biloba")
```

### Plot the locations of tree species "Cercis occidentalis" in San Francisco County
```{r}
sftreeplot("Cercis occidentalis")
```


