---
title: "Chemical Compositions"
subtitle: Homework from Thursday lab week 6 on tables and plots
date: March 3 2022
author: Jessica Hunter
output: 
  html_document:
    toc: TRUE
---



## Libraries
#### This is a list of the libraries used.
```{r, echo=TRUE,  results='hide', message=FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(kableExtra)
```

## Data
#### _This section lists the data that is loaded._
1. CondData.csv  
2. DepthData.csv

## Functions
#### _This section is a list of the functions used._
* The first function rounds the conductivity data to the nearest 10 seconds so I can merge with the depth data later  
* The next function converts the date and time to date and time this function joins the two data together where there are no NA's and only exact matches. It also rounds the time to 30 minute and takes the mean of the date, temp, depth, and salinity.


```{r, message=FALSE, warning=FALSE, echo=TRUE}
condata <- read_csv(here("week_6","data6", "CondData.csv")) %>% # read in the data
  mutate(dateandtime = mdy_hms(depth), # convert to datetime format
         dateandtime = round_date(dateandtime, "10 seconds")) %>% # rounds to the nearest 10 seconds
  drop_na() %>%  # dropping missing data rows to remove na's
  select(!depth)  # this removes the depth column
```
## Data Visualization  
#### This section plots a continuous **time** series of **salinity** readings on **January 15, 2022**

```{r, message=FALSE, warning=FALSE, echo=TRUE}
depdata <- read_csv(here("week_6", "data6", "DepthData.csv")) %>% # reds in data
  mutate(dateandtime = ymd_hms(date)) # converts to time format

inner_join(depdata,condata) %>% # joins the two and omits the ones that don't match
  select(!date) %>% # this removes the original date column
  mutate(timebyminute = round_date(dateandtime, "30 minutes")) %>% # this rounds by the half hour
  group_by(timebyminute) %>% # grouping variable is time
  summarise_at(vars(c("dateandtime","TempInSitu","Depth","SalinityInSitu_1pCal")), 
               list(mean = mean)) %>% # takes mean of each column listed
  mutate(Time = format(timebyminute, format = "%H:%M")) %>% # this changes to hour and minute
  select(Time, "Salinity" = SalinityInSitu_1pCal_mean)  %>% # only selects time and salinity columns
   kbl(caption = "Salinity by the half-hour") %>% # makes a table
   kable_classic("striped") %>% # theme
   row_spec(c(0), bold = TRUE, color = "white", background = "black") %>% # changes top row to black
     kable_styling(full_width = FALSE) #to not make whole width of page
```



```{r, message=FALSE, warning=FALSE, echo=TRUE}
# this function plots the salinity by minute
depdata %>% # reads in data
  inner_join(condata) %>% # joins two files and removes non matching
  rename("Salinity" = SalinityInSitu_1pCal)%>% # renames columns
  filter(Salinity>30)%>% # removes outlier
 # mutate(Time = hm(Time)) %>% 
# view(dep_con_data)
  ggplot(mapping = aes(x= date, # this maps the x and y
                       y =Salinity)) +
  geom_point(alpha = .1, color = "purple", # adds data point
             size = 2)+
  geom_smooth(color = "orange", 
              size = 2) + # adds line through data points
    labs(title = "Salinity over Time", # adds labels
       subtitle = "Time average by minute for date January 15, 2022",
       x = "Time (min)",
       y = "Salinity (ppm)") +
  theme(plot.title = element_text(size = 20), # apply title
        axis.title = element_text(size = 15), # axis title
        panel.background = element_rect(fill = "grey"), # colors background
        plot.background = element_rect(color = "grey")) +
  scale_x_datetime(date_breaks = "1 hour", date_labels = "%H") # adds manual tick marks
```


