---
title: "w6thurs"
author: "Jessica Hunter"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../output6/", warning = FALSE)
## two dots means go up a level and then go to output
```


```{r, echo=TRUE}
# install.packages("kableExtra")
# install.packages("ymlthis")
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
```

```{r penguin1, out.width = "50%"}
ggplot(penguins,
       aes(x=bill_length_mm,
           y = bill_depth_mm))+
  geom_tile()
```

out for output and fig for figure  

out.width="70%" (makes it 70% of the width) -->  
fig.width=3 (makes the width 3 inches) -->  
fig.height=4 (makes the width 4 inches) -->  
fig.asp=1.2 (changes the aspect ratio to 1.2) -->  
fig.align="center" -->  
fig.path in the the top so apply to whole doc but cant use here
r label,  
fig path cant use here function :(  
have to change label names for each one or it will keep saving over  
echo hides code
{r label, option1=value}

```{r penguin2, fig.align='center',fig.width=3, fig.height=4, fig.asp=1.2, fig.cap="this is a plot for penguins"}
ggplot(penguins,
       aes(x=bill_length_mm,
           y = bill_depth_mm))+
  geom_point()
```


handwrite table
can make tables interactive
twitter competition for r tables
google different tables  
[link to how to create tables](https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/)  
[link for themes of outputs](https://www.datadreaming.org/post/r-markdown-theme-gallery/)

you can save to output github_document to view easier in github
pdf doc takes 5 minutes to make file
look in YAMLthis is addins tab >



```{r}
penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() %>%  # make it a kable table
    kable_classic()%>%
  row_spec(2, bold = T, color = "white", background = "red") %>% 
    kable_styling(full_width = FALSE) #to not make it super wide of the whole width


```

```{r}
library(memer)

```

