---
title: "Rmarkdown Notes"
output: 
  html_notebook
    toc: TRUE
    toc_float: TRUE
    theme: darkly
---

This is the notes for class week6a
include=false - dont want to show code ot results
echo = false - shows results but not code
message
warning
fig.cap = "jflj;dj"




#YAML -yet another markdown language
if not installed yet install:
tinytex and
rmarkdown

global - this section sets for all moving forward
```{r setup, include = FALSE}
knit::opts_chunks$set(warning = FALSE,
                      echo = TRUE)
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

