---
title: "w12_words"
author: "Jessica Hunter"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
regex is regular expression
```{r eval=FALSE, include=FALSE}
install.packages("tidytext")
install.packages("wordcloud2")
install.packages("janeaustenr")
install.packages("stopwords")
```
library()


```{r}
library(here)
library(tidyverse)
library(tidytext)
library(wordcloud2)
library(janeaustenr)
```


## stringr

paste("first word", "second word", sep = "-") manioulation
can use column names but then remove ""  
paste0() means no space between  
can out in middle end etc as long as in between  
shapes <- c("Square", "Circle", "Triangle")
paste("My favorite shape is a", shapes)  


```{r}
shapes <- c("Square", "Circle", "Triangle")
paste("My favorite shape is a", shapes)
str_length(shapes) # how many letters are in each word?
seq_data<-c("ATCCCGTC")
str_sub(seq_data, start = 2, end = 4) # extract the 2nd to 4th AA
seq_data<-c("ATCCCGTC")
str_sub(seq_data, start = 2, end = 4) # extract the 2nd to 4th AA
str_sub(seq_data, start = 3, end = 3) <- "A" # add an A in the 3rd position
seq_data

#You can also duplicate patterns in your strings. Here I am duplicating it 2 and 3 times

str_dup(seq_data, times = c(2, 3)) # times is the number of times to duplicate each string

```



```{r}
badtreatments<-c("High", " High", "High ", "Low", "Low")
badtreatments
str_rim(badtreatments) #string trim
str_trim(badtreatments, side = "left") # this removes left
str_pad()
str_pad(badtreatments, 5, side = "right") # add a white space to the right side after the 5th character
str_pad(badtreatments, 5, side = "right", pad = "1") # add a 1 to the right side after the 5th character
#string to upper and loweer
x<-"I love R!"
str_to_upper(x)
str_to_lower(x)
str_to_title(x) #adds cap to first letter in each word
data<-c("AAA", "TATA", "CTAG", "GCTT")

# find all the strings with an A
str_view(data, pattern = "A")
str_detect(data, pattern = "A")

str_detect(data, pattern = "AT")
str_locate(data, pattern = "AT")

```

stringr cheat sheet [link](http://edrub.in/CheatSheets/cheatSheetStringr.pdf)  


## regex  

  Metacharacters
  Sequences
  Quantifiers
  Character classes
  POSIX character classes (Portable Operating System Interface)

these are used in r so have to use 2 backslashes to exscape the function  
. \ | ( ) [ { $ * + ?

### sequences

```{r}
vals<-c("a.b", "b.c","c.d")
#string, pattern, replace
str_replace(vals, "\\.", " ") # \\ escapes the function of the dot
vals<-c("a.b.c", "b.c.d","c.d.e")
#string, pattern, replace
str_replace(vals, "\\.", " ") # only removes first one  
str_replace_all(vals, "\\.", " ") # to do all of them and not only the first one


```

### sequences
 as the name suggests refers to the sequences of characters which can match. We have shorthand versions (or anchors) for commonly used sequences in R: 
 
```{r}
val2<-c("test 123", "test 456", "test")
str_subset(val2, "\\d") #subsets the things that fit the pattern

```

### character class
A character class or character set is a list of characters enclosed by square brackets [ ]. Character sets are used to match only one of the different characters. For example, the regex character class [aA] matches any lower case letter a or any upper case letter A. 
 a carrot means anything other than
 
```{r}
str_count(val2, "[aeiou]")

## [1] 1 1 1

# count any digit
str_count(val2, "[0-9]")
```
 
 
### quantifiers

**makes sure to use \\**

ˆ
```{r}
strings<-c("550-153-7578",
         "banana",
         "435.114.7586",
         "home: 672-442-6739")
phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
test<-str_subset(strings,phone)
test
# Which strings contain phone numbers?
str_detect(strings, phone)

```
Let's clean it up. Lets replace all the "." with "-" and extract only the numbers (leaving the letters behind). Remove any extra white space. You can use a sequence of pipes.

I wanted to extract the digits but then I need to add the dashes back in

```{r}
str_replace_all(test,"\\.","-")

test %>%
  str_replace_all(pattern = "\\.", replacement = "-") %>% # replace periods with -
  str_replace_all(pattern = "[a-zA-Z]|\\:", replacement = "") %>% # remove all the things we don't want
  str_trim() # trim the white space
```

## tidytest

we will analyze the books by jane austin
```{r}
head(austen_books())
```
 add a line for line and chapter
```{r}
original_books <- austen_books() %>% # get all of Jane Austen's books
  group_by(book) %>% # so i can oull each ch
  mutate(line = row_number(), # find every line
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", # count the chapters (starts with the word chapter followed by a digit or roman numeral)
                                                 ignore_case = TRUE)))) %>% #ignore lower or uppercase
  ungroup() # ungroup it so we have a dataframe again
# don't try to view the entire thing... its >73000 lines...
head(original_books)
```

clean this so that there is only one word per row so its tidy. In tidytext each word is refered to as a token. The function to unnest the data so that its only one word per row is unnest_tokens().

```{r}
tidy_books <- original_books %>%
  unnest_tokens(output = word, input = text) # add a column named word, with the input as the text column
head(tidy_books) # there are now >725,000 rows. Don't view the entire thing!
```
  Words that are common and don't really have important meaning (e.g. "and","by","therefore"...). These are called stopwords. We can use the function "get_stopwords()" to essentially remove these words from our dataframe. (This function is essentially just a dataframe of unnecessary words)
```{r}
#see an example of all the stopwords
head(get_stopwords())
cleaned_books <- tidy_books %>%
  anti_join(get_stopwords()) # dataframe without the stopwords

## Joining, by = "word"

head(cleaned_books)

```
  Can create a vector that includes the keywords that i'm looking for
  
```{r}
cleaned_books %>%
  count(word, sort = TRUE)
```
  There are many ways that we can now analyze this tidy dataset of text. One example is we could do a sentiment analysis (how many positive and negative words) using get_sentiments(). 
```{r}
sent_word_counts <- tidy_books %>%
  inner_join(get_sentiments()) %>% # only keep pos or negative words
  count(word, sentiment, sort = TRUE) # count them
```

Now, think about how we could do this with science? Instead of get_sentiments(), you could use an inner_join with a vector of keywords that you are searching for.   

lets plot the words


```{r}
sent_word_counts %>%
  filter(n > 150) %>% # take only if there are over 150 instances of it
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>% # add a column where if the word is negative make the count negative
  mutate(word = reorder(word, n)) %>% # sort it so it gows from largest to smallest
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col() +
  coord_flip() +
  labs(y = "Contribution to sentiment")
```

word cloud

```{r}
words<-cleaned_books %>%
  count(word) %>% # count all the words
  arrange(desc(n))%>% # sort the words
  slice(1:100) #take the top 100
wordcloud2(words, shape = 'triangle', size=0.3) # make a wordcloud out of the top 100 words
```

  ## r package
  
  link for [gg in real life](https://jnolis.com/blog/introducing_ggirl/)
  
  
  
  
 