---
title: "Iterative loops"
author: "Jessica Hunter"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## library

```{r}
library(tidyverse)
library(here)
```


## load data

```{r}

```
## notes

[purrcheatsheet](https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf)
[r4ds chapter](https://r4ds.had.co.nz/iteration.html)
[advanced r](https://adv-r.hadley.nz/functionals.html)


all file names have to be the same, one can't have . and one has _
you can use file choose in shiny app

### for loops

for(index for sequencing){
command to repeat
}
build from inside out
```{r}
# simple for loop
print(paste("This year is", 2000))

years<- c(2015:2021)
for (i in years){
  print(paste("This year is", i))
}
```

create dataframe
always use lnegth() and ncol because the numbers might change in the future
common mistake is to forget to include value on the left side have tell it which row it's in 1st row in 1st year

```{r}
year_data<- data.frame(matrix(ncol = 2, nrow = length(years)))
colnames(year_data)<- c("year", "year_name")

year_data

for(i in 1:length(years)){
  year_data$year_name[i] <-paste("This year is", years[i] )
  year_data$year[i]<-years[i] # loop over year

  }

```

test it out

```{r}
testdata<- read_csv(here("week_13", "data13", "011721_CT354_1pcal.csv" ))
glimpse(testdata)
# point to the location on the computer of the folder
CondPath<-here("week_13", "data13", "cond_data")
# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# you can use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath,pattern = ".csv")
files
```

 pre-allocate space
```{r}

# make an empty dataframe that has one row for each file and 3 columns
cond_data<-data.frame(matrix(nrow = length(files), ncol = 3))
# give the dataframe column names
colnames(cond_data)<-c("filename","mean_temp", "mean_sal")
cond_data
```
Turn into for loop
add in the raw data over the loop
add columns
add in means


test first one
```{r}
raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works
glimpse(raw_data)
class(raw_data)

for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
}
```

### purr
purr faster
also there is a for each package that runs loops in paralell

map functions

The pattern of looping over a vector, doing something to each element and saving the results is so common that the purrr package provides a family of functions to do it for you. There is one function for each type of output:

  map() makes a list.
   map_lgl() makes a logical vector.
   map_int() makes an integer vector.
  map_dbl() makes a double vector.
  map_chr() makes a character vector.
  map_df() makes a dataframe

Each function takes a vector as input, applies a function to each piece, and then returns a new vector that’s the same length (and has the same names) as the input.


3 ways in map function
do it straight
do it create own function
do it with function already in r

tilda in front and put a dot infront of the variable you are changing when doing multiple things with a function
```{r}
1:10 %>% 
  map(rnorm, n=15) %>% # 15 random number do it 10 times
  map_dbl(mean)
  
1:10 %>% 
  map(function(x) rnorm(15, x)) %>% 
  map_dbl(mean)

1:10 %>% 
  map(~ rnorm(15, .x)) %>% 
  map_dbl(mean)

```

bring in files
purr turns all of my files into a dataframe

```{r}
#Reminder: find the files

# point to the location on the computer of the folder
CondPath<-here("week_13", "data13", "cond_data")
files <- dir(path = CondPath,pattern = ".csv")
files



#Or, we can get the full file names in one less step by doing this...

files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files

## this brings in the data and puts it on top of each other maps on top of csv
#Next, read in the files using map instead of a for loop while retaining the filename as a column.

data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>%  # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```


## awesome r package
[aanatogram](https://r-charts.com/miscellaneous/gganatogram/)






