---
title: "models"
author: "Jessica Hunter"
date: "4/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE)
```


```{r eval=FALSE, include=FALSE}
#install.packages('modelsummary')
install.packages('tidymodels')
install.packages('broom')
install.packages('flextable')
#install.packages('peformance')
library(performance)
```


```{r, results='hide', message=FALSE}
library(tidyverse)
library(here)
library(palmerpenguins)
library(broom)
library(performance) 
library(modelsummary)
library(tidymodels)
```
linear model
mod<-lm(y~x, data = df)

lm = linear model, y = dependent variable, x = independent variable(s), df = dataframe.

You read this as y is a function of x

Multiple regression
mod<-lm(y~x1 + x2, data = df)

Interaction term
mod<-lm(y~x1*x2, data = df) the * will compute x1+x2+x1:x2
```{r}
Peng_mod<-lm(bill_length_mm ~ bill_depth_mm*species, data = penguins)
check_model(Peng_mod) # check assumptions of an lm model


```

anova() gives type 3
Anova with capital you can chooses type a or 2

## broom

tidy gives summary coeeficients

glance give stats interpretation of whole model, calc fitted value at each x

~tilda in front of function in map function allows you to change one variable


```{r}
anova(Peng_mod)
summary(Peng_mod)

# Tidy coefficients
coeffs<-tidy(Peng_mod) # just put tidy() around it
knitr::kable(coeffs)

#glance extracts R-squared, AICs, etc of the model

# tidy r2, etc
results<-glance(Peng_mod) 
knitr::kable(results)

# augment add residuals and predicted values to your original data and requires that you put both the model and data

# tidy residuals, etc
resid_fitted<-augment(Peng_mod)
knitr::kable(head(resid_fitted))
```

## Results in modelsummary

{modelsummary} creates tables and plots to summarize statistical models and data in R.

modelsummary includes two families of functions:

Model Summary
modelsummary: Regression tables with side-by-side models.
modelsummary_wide: Regression tables for categorical response models or grouped coefficients.
modelplot: Coefficient plots.

Data Summary
datasummary: Powerful tool to create (multi-level) cross-tabs and data summaries.
datasummary_balance: Balance tables with subgroup statistics and difference in means (aka “Table 1”).
datasummary_correlation: Correlation tables.
datasummary_skim: Quick overview (“skim”) of a dataset.
datasummary_df: Turn dataframes into nice tables with titles, notes, etc. 
Export summary tables to word, markdown, or tex document. You can also modify the tables to make them pub quality.

Let's compare the Peng_mod with one that does not have species as an interaction term.

```{r}
# New model
Peng_mod_noX<-lm(bill_length_mm ~ bill_depth_mm, data = penguins)
#Make a list of models and name them
models<-list("Model with interaction" = Peng_mod,
             "Model with no interaction" = Peng_mod_noX)
#Save the results as a .docx
modelsummary(models, output = here("week_13","output13","table.docx"))
```

## 

Canned coefficient modelplots
compare both models with and without c
```{r}
#install.packages("wesanderson")
library(wesanderson)
modelplot(models) +
    labs(x = 'Coefficients', 
         y = 'Term names') +
    scale_color_manual(values = wes_palette('Darjeeling1'))
```

Many models with purrr, dplyr, and broom

Let's say you want to plot and compare lots of different models at the same time and view the results. For example, instead of using species as an interaction term, let's make an individual model for every species.

We can essentially make a set of lists that have each dataset that we want to model and use the map functions to run the same model to every dataset. We will test it step by step

First, let's call the penguin data and create a list for the data by each species. We do this using nest(). We are going to nest the data by species. 

```{r}
 models<- penguins %>%
  ungroup()%>% # the penguin data are grouped so we need to ungroup them
    nest(-species) %>%  # nest all the data by species
  mutate(fit = map(data, ~lm(bill_length_mm~body_mass_g, data = .)))
models
```


```{r}
#View the results. 
# First, let's mutate the models list so that we have a tidy coefficient dataframe (using tidy()) and a tidy model results dataframe (using glance())

 results<-models %>%
   mutate(coeffs = map(fit, tidy), # look at the coefficients
          modelresults = map(fit, glance)) %>% # R2 and others 
   select(species, coeffs, modelresults) %>% # only keep the results
   unnest() # put it back in a dataframe and specify which columns to unnest

view(results) # view the results
```

## Other very common stats packages

stats: General (lm)and generalized (glm) linear models (already loaded with base R)  
lmer : mixed effects models
   lmerTest' : getting results from lmer
  nlme : non-linear mixed effects models  
  mgcv, gam : generalized additive models
  brms, rstan, and many more : Bayesian modeling  
  lavaan, peicewiseSEM : Structural Equation Models
  rpart, randomForest, xgboost, and more : Machine learning models


## tidymodels


```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>% # linear model
   fit(bill_length_mm ~ bill_depth_mm*species, data = penguins) %>% # add model fit
  tidy() %>% 
    ggplot()+
    geom_point(aes(x = term, y = estimate))+
    geom_errorbar(aes(x = term, ymin = estimate-std.error,
                      ymax = estimate+std.error), width = 0.1 )+
  coord_flip()
lm_mod
```

