---
title: "Week 13 homework"
author: "Jessica Hunter"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Intro

This homework script will create two loops to analyze more than one file at a time and place results in a new dataframe

## Libraries

These are the libraries needed to run the code

```{r, results='hide'}
library(tidyverse)
library(here)
```


## Data

This is my data used by naming the path

```{r, results='hide'}
HWpath<-here("week_13", "data13", "homework") # creating a path

files <- dir(path = HWpath,pattern = ".csv") # not using full name 

files2 <- dir(path = HWpath,pattern = ".csv", full.names = TRUE) # using full name for map function
```


## Analyze

### Using a for loop to summarize mean and standard deviation for light and temperature of each file

```{r}
# create an empty data frame to hold new entries

hw<- data.frame(matrix(nrow = length(files), ncol = 5))
colnames(hw)<-c("file_name", "mean_temp_C", "stddev_temp_C", "mean_light_lux", "stddev_light_lux")

for(i in 1:length(files)){ # use length of files in case number changes in future
  rawdata<- read_csv(here(HWpath,files[i])) # read in data
  hw$file_name[i]<-files[i] # paste new entry new dataframe
  hw$mean_temp_C[i]<- mean(rawdata$Temp.C, na.rm = TRUE) # calculate mean and paste entry
  hw$stddev_temp_C[i]<- sd(rawdata$Temp.C, na.rm = TRUE) # calculate standard deviation and paste entry
  hw$mean_light_lux[i]<- mean(rawdata$Intensity.lux, na.rm = TRUE) # calculate mean and paste entry
  hw$stddev_light_lux[i]<- sd(rawdata$Intensity.lux, na.rm = TRUE) # calculate standard deviation and paste entry
}

knitr::kable(hw, caption = "Mean and Standard Deviation of Temperature and Light") # view table

```

### Using map to summarize mean and standard deviation for light and temperature of each file

```{r}
summarymap<- files2 %>% # create new dataframe
  set_names() %>% # names the vector
  map_df(read_csv,.id = "filename") %>% #reads in file, requires full name
  mutate(filename=gsub(".csv","",basename(filename))) %>% # removes full name and path
  group_by(filename) %>% # groups for next function
  summarise(mean_temp_C = mean(Temp.C, na.rm = TRUE), # calculate mean
            stddev_temp_C = sd(Temp.C, na.rm = TRUE), # calculate standard deviation
            mean_light_lux = mean(Intensity.lux, na.rm = TRUE), #calculate mean
            stddev_light_lux = sd(Intensity.lux, na.rm = TRUE)) # calculate standard deviation

  knitr::kable(summarymap, caption = "Mean and Standard Deviation of Temperature and Light") # view table
  
```





