---
title: "R Notebook"
output: html_notebook
---



```{r}

```{r}
sf_common_count <- sf_tree_names %>% 
  #select(common) %>% 
  group_by(common, .add = TRUE) %>% 
  add_count()
#filter(unique(sf_common_count$common))
#%>% 

#unique(common)
```



```{r}
testvector<- unique.data.frame(sf_common_count)
test2vector<-testvector$common
test3vector<-testvector$n

nameswitch<-function(input){
  for (i in seq_along(test2vector))
  print(switch(i,test3vector[i]))
}
if "Victorian Box" %in% test2vector print("true")

vectornumber<- function(input1){
  output<-test2vector
  return(output)
  
}
nameswitch("Victorian Box")
test2vector[1]

practice1<- function(y){
  for(y in test2vector) 
  print(y)
}
practice1("Victorian Box")

function my_idea
A = [ 1 2 5 7 ];
n = 8;
M = [ 3 4 6 9];
for i= 1:n
    switch i
        case num2cell(A)
            disp('suri');
        case num2cell(M)
            disp('kumar')
        otherwise
            disp('error')
    end
end
end

  switch(input,test3vector}
input<-
)
```






```{r}
for(i in 1:5) print(1:i)
i
for(n in c(2,5,10,20,50)) {
   x <- stats::rnorm(n)
   cat(n, ": ", sum(x^2), "\n", sep = "")
}
f <- factor(sample(letters[1:5], 10, replace = TRUE))
for(i in unique(f)) print(i)
i
1
f
n
1
for("Victorian Box" in sf_tree_names$common) print(sf_tree_names$scientific)
switch
c2s<- function(input) {
  switch(input,)
}
```



```{r}
scientificname <- function(input){
  if {{input}} %in% c("common", "othercommon"){
  print(testvector$scientific)
  }
  output<-testvector %>% 
    filter(common=={{input}} | othercommon=={{input}}) %>% 
    select(scientific)
  return(output$scientific)
}


hi<- testvector %>% 
  filter(common=="Victorian Box" | othercommon=="Victorian Box") %>% 
  select(scientific)
print(hi$scientific)
hi
 
test<- if ("Victorian Box" %in% c("common", "othercommon")) {
  print(scientific)
}
  #select(scientific)
  print(test$scientific)
  
nameswitch<- function(name,type){
  switch(type,
         scientific = blah
         common = blahblah
           )
}

```



```{r}
common2scientific<-function(data = sf_tree_names, input){
  output<-if((input %in% c("common","othercommon")
    return(print({{sf_tree_names$scientificname}}))
    return("check spelling and try again")))}
```


```{r}

#%>% 
filter(common=="Victorian Box")
sf_trees_clean %>% 
  ifelse(sf_trees_clean$common =="Victorian Box",sf_trees_clean$scientific,"check spelling")
```


## map plot that works


example 
```{r}
maptrees<- sf_trees_clean %>% 
  filter(scientific =="Acacia melanoxylon") %>% 
  ggplot(aes(x=longitude,y=latitude))+
  geom_point()+
   geom_polygon(data = sanfran,
                aes(y = lat,
                    x = long,
                    group = group,
                    alpha = 0.2))+
  coord_map(projection = "mercator",
             ylim = c(37.69,37.825),
             xlim = c(-122.5, -122.3))+
  theme(legend.position = "none")
#+
#  ggmap(sfmap,zoom=17)
maptrees

```



example works
```{r}
geocode("San Francisco County")
sfdataframe<-data.frame(lon= -122.4194, lat= 37.77493)
gmap1<-get_map(sfdataframe, zoom = 13, maptype = "satellite")
ggmap(gmap1)+
  geom_point(data = maptrees, 
             aes(x=longitude,y=latitude,
                 color = scientific))
  

##############################################
# sfmap<-get_map("San Francisco")
# ggmap(sfmap,zoom=17)+
#   geom_point(data = maptrees,aes(x=longitude,y=latitude, scientific = "red"))
# 
# sfmap<-get_map("San Francisco")
# maptrees<- sf_trees_clean %>%
#   filter(scientific =="Acacia melanoxylon")
# # %>%
#   ggmap(sfmap,zoom=17)+
#   #ggplot(aes(x=longitude,y=latitude))+
#   geom_point(aes(x=longitude,y=latitude))
  
  # geom_polygon(data = sanfran,
  #               aes(y = lat,
  #                   x = long,
  #                   group = group,
  #                   alpha = 0.2))+
  # coord_map(projection = "mercator",
  #            ylim = c(37.69,37.825),
  #            xlim = c(-122.5, -122.3))+
  # theme(legend.position = "none")+
  
maptrees
sfmap<-get_map("San Francisco")
ggmap(sfmap,zoom=17)
```





```{r}
treeplot<- function(data=maptrees, name){
  maptrees<-sf_trees_clean %>% 
    filter(scientific=="{{name}}")
  #return(maptrees)
}
test3<- "hi"
test2<-"Arbutus"
test<- c("Arbutus","Ginkgo biloba")
if (test %in% sf_trees_clean$scientific) {
  print("yes")
}else{
  print(test2)
}

  print(test[2])
test
print(test[2])


############## this one works
geocode("San Francisco County")
maptrees<- sf_trees_clean %>% 
  filter(scientific =="Acacia melanoxylon")
sfdataframe<-data.frame(lon= -122.4194, lat= 37.77493)
gmap1<-get_map(sfdataframe, zoom = 13, maptype = "satellite")
ggmap(gmap1)+
  geom_point(data=maptrees,
             aes(x=longitude,y=latitude,
                 color = scientific))
  
```



trying with ggmaps

```{r}
sfdataframe<-data.frame(lon= -122.4194, lat= 37.77493)
gmap1<-get_map(sfdataframe, zoom = 13, maptype = "satellite")
ggmap(gmap1)
```



### trying to use switch not working

```{r}
treenames<- unique.data.frame(sf_tree_names)
practice5<- function(samplename, type, data=treenames){
  switch(type,
         scientific= {
           if (samplename %in% treenames$scientific){
    efgh<-filter(treenames, scientific=={{samplename}}) %>% 
    print()
         }
         common = {if (samplename %in% treenames$common){
    ijkl<-filter(treenames, common=={{samplename}}) %>% 
    print()}
         
         
           )
  if (samplename %in% treenames$scientific){
    efgh<-filter(treenames, scientific=={{samplename}}) %>% 
    print()
  } else{
    print("no")
  }
}

practice5("Ginkgo biloba")
practice5("Cercis occidentalis")
```



```{r}
practice7<-function(samplename,type, data=treenames){
if (samplename %in% treenames$type){
    filter(treenames, {{type}}=={{samplename}}) %>% 
    print()
} else {
  print("error")
}
}
practice7("Ginkgo biloba",scientific)
```



