---
title: "San Francisco Trees"
author: "Jessica Hunter"
date: "3/8/2022"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro



## Libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapproj)
library(mapdata)
library(ggplot2)
```


## load data
sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

```{r}
sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

## Functions

```{r}
# to make a map of san francisco
sanfran <- map_data("county") %>% 
 filter(subregion == "san francisco")

# creates the columns into the scientific or common name
sf_trees_clean <-sf_trees %>% 
  select(species, latitude, longitude) %>% 
  separate(col = species,
           into = c("scientific","common"),
           sep = " :: ") %>% 
  filter(common != "")
 
 view(sf_trees_clean)
  
sf_common <- sf_trees_clean %>% 
  select(common, latitude, longitude) %>% 
  filter(common != "") %>% 
  group_by(common, .add = TRUE) %>% 
  add_count() %>% 
  filter(n > 150)
  #### cant get to work right
  
view(sf_common)
amelm <- sf_common %>% 
  filter(common == c("Victorian Box"))
view(amelm)

amelm<- amelm %>%
  filter(latitude > 37.65,
         longitude > -122.5) %>%
  droplevels()


unique(amelm$latitude)
cityfilter<-sf_common %>% 
  filter(longitude < min(long))

ggplot()+
   geom_polygon(data = sanfran,
                aes(y = lat,
                    x = long,
                    group = group))+
  geom_point(data = amelm,
           aes(x = longitude,
              y = latitude,
             color = common))+
   coord_map(projection = "mercator",
             ylim = c(37.69,37.825),
             xlim = c(-122.5, -122.3))+
  theme(legend.position = "none")


ggplot()+
   geom_polygon(data = sanfran,
                aes(x = long,
                    y = lat,
                    group = group))+
  geom_point(data = sf_common,
             aes(x = longitude +0.025,
                 y = latitude,
                 color = common))+
  coord_map(ylim = c(37.69,37.825),
            xlim = c(-122.5, -122.3))+
  theme(legend.position = "none")

```


```{r}
sf_common <- sf_trees_clean %>% 
  select(common, latitude, longitude) %>% 
  filter(common != "") %>% 
  group_by(common) %>% 
  add_count()
```

new function
```{r}

  

treeminmaxlatlong <- sf_common %>% 
  filter(latitude < 37.81522 | latitude > 37.70062,
         longitude < -122.3208 | longitude > -122.4696) %>% 
  group_by(latitude) %>% 
  summarize(longmin = min(longitude, na.rm = TRUE),
            longmax = max(longitude, na.rm = TRUE))

avglong <- sf_common %>% 
  mutate(avgminlong = )



sfminmaxlat <- sanfran %>% 
  group_by(long) %>% 
  summarise(latmin = min(lat, na.rm = TRUE),
            latmax = max(lat, na.rm = TRUE))

```


```{r}

  
  # add_count()

sfminmaxlong <- sanfran %>% 
  group_by(lat) %>% 
  summarize(longmin =  min(long, na.rm = TRUE),
            longmax = max(long, na.rm = TRUE))  %>% 
  mutate(abc = c("a","b","c","d","e","f","g","h","i"))


x <- sfminmaxlong$lat
a <- x[1]
b <- x[2]
c <- x[3]
d <- x[4]
e <- x[5]
f <- x[6]
g <- x[7]
h <- x[8]
i <- x[9]
x
library(dplyr)
sfmap <- sfminmaxlong %>% 
  mutate(latitude = lat)


sftrees <- sf_common %>% 
  filter(common != "") %>% 
  filter(latitude < 45 & latitude>37.6) %>% 
  group_by(common) %>% 
  mutate(latitude, avglat =
           if_else(between(latitude,37.7006,mean(37.70062,37.71781)),37.71781,
if_else(between(latitude,mean(37.70062,37.71781),mean(37.71781,37.72927)),37.72927,
if_else(between(latitude,mean(37.71781,37.72927),mean(37.72927,37.73500)),37.73500,
if_else(between(latitude,mean(37.72927,37.73500),mean(37.73500,37.76365)),37.76365,
if_else(between(latitude,mean(37.73500,37.76365),mean(37.76365,37.78657)),37.78657,
if_else(between(latitude,mean(37.76365,37.78657),mean(37.78657,37.79230)),37.79230,
if_else(between(latitude,mean(37.78657,37.79230),mean(37.79230,37.80376)),37.80376,
if_else(between(latitude,mean(37.79230,37.80376),37.81522),37.81522,0)))))))))

ggplot(sftrees, aes(x=longitude,
                    y = latitude)) +
  geom_point()
  
 
#c,d),mean(d,e))),"e",
 #  d,e),mean(e,f))),"f",
   if_else(between(latitude,mean(e,f),mean(f,g)),"g",
   if_else(between(latitude,mean(f,g),mean(g,h)),"h",
   if_else(between(latitude,mean(g,h),i),"i","NA")))
  


sftrees <- amelm %>% 
  filter(common != "") %>% 
  group_by(common) %>% 
  mutate(avglat = mean(a,b),"b","z")
```

```{r}
sftrees <- amelm %>% 
  filter(common != "") %>% 
  #group_by(common) %>% 
  mutate(
    avglat = if_else(amelm$latitude > mean(37.70062,37.71781),37.71781,
if_else(amelm$latitude > mean(37.71781,37.72927),37.72927,
if_else(amelm$latitude > mean(37.72927,37.73500),37.73500,
if_else(latitude > mean(37.73500,37.76365),37.76365,
if_else(latitude > mean(37.76365,37.78657),37.78657,
if_else(latitude > mean(37.78657,37.79230),37.79230,
if_else(latitude > mean(37.79230,37.80376),37.80376,
if_else(latitude > 37.80376,37.81522,0)))))))))

```

